---
title: "R Notebook"
output: html_notebook
---
```{r}
library(dplyr)
library(tidyr)

```
Purpose: Created to compare the output/results from MAGeCK to those obtained from JACKS. Stopped working on it b/c JACKS measures gRNA efficacy while MAGeCK is focused on gRNA abundance. While possible to compare the algorithm's performance, you'd need to look into a common outcome metric.

# Import data
```{r import stmn2 MAGeCK sgrna and gene summaries and jacks stmn2 results}
stmn2_mageck_sgrna_df <- read.table("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/data_processed/rra-High-Low-STMN2.sgrna_summary.txt", sep = "\t", header = TRUE)
stmn2_mageck_gene_df <- read.table("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/data_processed/rra-High-Low-STMN2.gene_summary.txt", sep = "\t", header = TRUE)

stmn2_jacks_sgrna_df <- read.table("/Users/Claire/Downloads/JACKS-master/jacks_2018_paper/results_jacks/stmn2_jacks_results_3_samples/stmn_2_jacks_results_grna_JACKS_results.txt", sep = "\t", header = TRUE)
stmn2_jacks_sgrna_lfc <- read.table("/Users/Claire/Downloads/JACKS-master/jacks_2018_paper/results_jacks/stmn2_jacks_results_3_samples/stmn_2_jacks_results_logfoldchange_means.txt", sep = "\t", header = TRUE)
stmn2_jacks_gene_df <- read.table("/Users/Claire/Downloads/JACKS-master/jacks_2018_paper/results_jacks/stmn2_jacks_results_3_samples/stmn_2_jacks_results_gene_JACKS_results.txt", sep = "\t", header = TRUE)
stmn2_jacks_gene_sd <- read.table("/Users/Claire/Downloads/JACKS-master/jacks_2018_paper/results_jacks/stmn2_jacks_results_3_samples/stmn_2_jacks_results_gene_std_JACKS_results.txt", sep = "\t", header = TRUE)
stmn2_jacks_gene_pval <- read.table("/Users/Claire/Downloads/JACKS-master/jacks_2018_paper/results_jacks/stmn2_jacks_results_3_samples/stmn_2_jacks_results_gene_pval_JACKS_results.txt", sep = "\t", header = TRUE)


# Merge JACKS sgRNA efficacy file with sgRNA LFC 
# sgrna_merged <- dplyr::full_join(stmn2_jacks_sgrna_df, stmn2_jacks_sgrna_lfc)

# Merge JACKS gene pval and sd files with JACKS gene results file
gene_merged <- full_join(stmn2_jacks_gene_pval, stmn2_jacks_gene_sd, by = "Gene", suffix = c("_pval", "_sd"))
stmn2_jacks_gene_df <- full_join(stmn2_jacks_gene_df, gene_merged)
```

# Are the counts in the MAGeCK output really the raw sgRNA counts?

# Test normalization methods with various datasets
```{r}
stmn2_mageck_sgrna_df
neg_ctrl_stmn2_sgrna_sep <- stmn2_mageck_sgrna_df %>%
  dplyr::select(sgrna, Gene, control_count, treatment_count) %>% 
  tidyr::separate(control_count, into = c("low1_rep1", "low2_rep2", "low3_rep3"), sep = "/") %>% 
  tidyr::separate(treatment_count, into = c("high1_rep1", "high2_rep2", "high3_rep3"), sep = "/") %>% 
  tidyr::pivot_longer(cols = starts_with(start_strings),
               names_to = "bin",
               values_to = "normalized_read_count")
```


### JACKS gene results (18746) have more obs than MAGeCK (18327) - figure out why JACKS have 419 more rows
```{r}
genes_not_in_stmn2_jacks_gene_df %>% 
  dplyr::filter(!Gene %in% stmn2_mageck_gene_df$id)

# Function to compare two vectors and print non-matching values
compare_vectors <- function(vector1, vector2) {
  # Ensure the inputs are vectors
  if (!is.vector(vector1) | !is.vector(vector2)) {
    stop("Both inputs need to be vectors")
  }
  
  # Find values in vector1 not in vector2
  not_in_vector2 <- setdiff(vector1, vector2)
  
  # Find values in vector2 not in vector1
  not_in_vector1 <- setdiff(vector2, vector1)
  
  # Print out the values that don't match
  if (length(not_in_vector2) > 0) {
    cat("Values in vector1 not in vector2:\n")
    print(not_in_vector2)
  } else {
    cat("All values in vector1 are present in vector2\n")
  }
  
  if (length(not_in_vector1) > 0) {
    cat("Values in vector2 not in vector1:\n")
    print(not_in_vector1)
  } else {
    cat("All values in vector2 are present in vector1\n")
  }
}

# set vector1 = JACKS gene output and vector2 = MAGeCK gene output
stmn_jacks <- stmn2_jacks_gene_df$Gene
stmn_mageck <- stmn2_mageck_gene_df$id

# Compare the vectors
compare_vectors(stmn_jacks, stmn_mageck)
```

# Rank plot
```{r MAGeCK rank plot of STMN2 screen data, echo=FALSE, warning=FALSE, include=FALSE}
# from rankplot.R file
library(tidyverse)
library(ggrepel)
rra.High.Low.STMN2.sgrna_summary <- read.delim("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/rra-High-Low-STMN2.sgrna_summary.txt")

# arrange rows by values in LFC column
rra.High.Low.STMN2.sgrna_summary <- rra.High.Low.STMN2.sgrna_summary %>%
  arrange(LFC)
rra.High.Low.STMN2.sgrna_summary <-rra.High.Low.STMN2.sgrna_summary %>% 
  mutate(
    zscore = scale(LFC), # scale calculates mean and sd of a vector then scales each element by subtracting mean and dividing by sd
    row = as.numeric(rownames(rra.High.Low.STMN2.sgrna_summary)) # create column to specify row number
  )

## Generate Rank Plot of all STMN2 sgRNA data ----
rra.High.Low.STMN2.sgrna_summary %>%
  ggplot(aes(row, zscore, label=Gene))+
  geom_point(data = rra.High.Low.STMN2.sgrna_summary %>% 
               filter(Gene!="negative_control"),
             color="black",
             alpha=0.7) +
  geom_point(data = rra.High.Low.STMN2.sgrna_summary %>% 
               filter(Gene=="negative_control"),
             color="grey",
             alpha=0.2) +
  geom_point(data = rra.High.Low.STMN2.sgrna_summary %>% 
               filter(Gene=="TARDBP"|
                        Gene=="STMN2"|
                        Gene=="RANBP1"),
             color="purple",
             alpha=0.7) +
  #geom_point(data = both %>% 
  #             filter((control_mean>=3000|
  #                       treat_mean>=3000),
  #                    Gene%in%transcription$Gene),
  #           color="blue",
  #           alpha=0.7)+
  #geom_point(data = both %>% 
  #             filter((control_mean>=3000|
  #                       treat_mean>=3000),
  #                    Gene%in%translation$Gene),
  #           color="green",
#           alpha=0.7)+
  geom_label_repel(data = rra.High.Low.STMN2.sgrna_summary %>% 
                     filter(Gene=="TARDBP"|
                              Gene=="STMN2"|
                              Gene=="RANBP1"),
                   fill="purple",
                   box.padding = 0.5,
                   segment.color="black",
                   alpha=0.6,
                   max.overlaps = Inf) +
  labs(
    title = "STMN2-mscarlet rank plot",
    caption = 
      "Grey: NT guides
    Purple: Genes of interest") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))



## Filter STMN2 points with means greater than 3000 ----
rra.High.Low.STMN2.sgrna_summary_filtered <- rra.High.Low.STMN2.sgrna_summary %>%
  filter(control_mean>=3000|
           treat_mean>=3000) %>% 
  arrange(LFC)

rra.High.Low.STMN2.sgrna_summary_filtered <- rra.High.Low.STMN2.sgrna_summary_filtered %>% 
  mutate(
    zscore = scale(LFC),
    row = as.numeric(rownames(rra.High.Low.STMN2.sgrna_summary_filtered))
  )

## Generate Rank Plot of STMN2 sgRNA data with a mean read count greater than or equal to 3000 ----
rra.High.Low.STMN2.sgrna_summary_filtered %>%
  ggplot(aes(row, zscore, label=Gene)) +
  geom_point(data = rra.High.Low.STMN2.sgrna_summary_filtered %>% 
               filter(Gene!="negative_control"),
             color="black",
             alpha=0.7) +
  geom_point(data = rra.High.Low.STMN2.sgrna_summary_filtered %>% 
               filter(Gene=="negative_control"),
             color="grey",
             alpha=0.2) +
  geom_point(data = rra.High.Low.STMN2.sgrna_summary_filtered %>% 
               filter(Gene=="TARDBP"|
                        Gene=="STMN2"
                      #  Gene=="RANBP1"
               ),
             color="purple",
             alpha=0.7) +
  #geom_point(data = rra.High.Low.STMN2.sgrna_summary_filtered %>% 
  #             filter(Gene%in%mitochondrion$id
  #             ),
  #           color="red",
  #           alpha=0.7)+
  #geom_point(data = both %>% 
  #             filter((control_mean>=3000|
  #                       treat_mean>=3000),
  #                    Gene%in%transcription$Gene),
  #           color="blue",
  #           alpha=0.7)+
#geom_point(data = both %>% 
#             filter((control_mean>=3000|
#                       treat_mean>=3000),
#                    Gene%in%translation$Gene),
#           color="green",
#           alpha=0.7)+
geom_label_repel(data = rra.High.Low.STMN2.sgrna_summary_filtered %>% 
                   filter(zscore < -5
                          # Gene=="RANBP1"
                          ),
                 box.padding = 0.5,
                 segment.color="black",
                 alpha=0.6,
                 max.overlaps = Inf) +
  labs(
    title = "STMN2-mScarlet rank plot filtered at 3000",
    caption = 
      "Grey: NT guides
    Purple: Genes of interest"
    ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))


## Rank stats ------
rra.High.Low.STMN2.sgrna_summary_filtered %>% 
  filter(Gene!="negative_control") %>% 
  summary

# mean and SD of negative control/non-targeting guides
rra.High.Low.STMN2.sgrna_summary_filtered %>% 
  filter(Gene=="negative_control") %>% 
  summarise(
    sd = sd(zscore),
    mean = mean(zscore)
  )

rra.High.Low.STMN2.sgrna_summary_filtered %>% 
  filter(Gene=="negative_control") %>% 
  summarise(
    sd = sd(zscore),
    mean = mean(zscore)
  )

# QUESTION: ask JH what this is for? ---------
up = -0.1179674+ 4*0.5991452
down = -0.1179674- 4*0.5991452

rra.High.Low.STMN2.sgrna_summary_filtered %>% 
  filter(Gene!="negative_control",
         zscore > up |
           zscore < down) %>% 
  summarise(n=n())


```


# Look at differences between non-targeting guides *within* each screen
```{r}
# Import STMN2 screen data
stmn2_mageck_sgrna_df <- read.table("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/data_processed/rra-High-Low-STMN2.sgrna_summary.txt", sep = "\t", header = TRUE)
stmn2_mageck_gene_df <- read.table("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/data_processed/rra-High-Low-STMN2.gene_summary.txt", sep = "\t", header = TRUE)

# Import TDP-43 screen data
tdp_mageck_sgrna_df <- read.table("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/tdp_i3N.sgrna_summary.txt", sep = "\t", header = TRUE)

# sgRNA hdr key: 
#   control_count: Normalized read counts in ctrl samples;
#   treatment_count: Normalized read counts in tx samples;
#   control_mean: Median read counts in ctrl samples;
#   treat_mean: Median read counts in tx samples
#QUESTION: how does the ctrl_count col differ from ctrl_mean col since 

# Plot null distribution of normalized non-targeting guides ----
# plot_density <- function(data, title) {
#   data %>%
#     mutate(log2_read_counts = log2(read_counts)) %>%
#     ggplot(aes(x = log2_read_counts, color = bin)) +
#     geom_density() +
#     labs(title = title, x = "Mean read counts (log2 scaled)", y = "Density") +
#     theme_minimal()
# }

# Create start_strings to select both high and low cols
start_strings <- c("high", "low")
# Separate the normalized read count cols
neg_ctrl_stmn2_sgrna_sep <- stmn2_mageck_sgrna_df %>%
  dplyr::select(sgrna, Gene, control_count, treatment_count) %>% 
  tidyr::separate(control_count, into = c("low1_rep1", "low2_rep2", "low3_rep3"), sep = "/") %>% 
  tidyr::separate(treatment_count, into = c("high1_rep1", "high2_rep2", "high3_rep3"), sep = "/") %>% 
  tidyr::pivot_longer(cols = starts_with(start_strings),
               names_to = "bin",
               values_to = "normalized_read_count") %>% 
  dplyr::filter(Gene=="negative_control")

# Create plots high/low FACS bins of normalized-log2 scaled sgRNA read counts
stmn2_high_neg_ctrl_sgrna_counts <- neg_ctrl_stmn2_sgrna_sep %>%
  dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
  dplyr::filter(startsWith(bin, "high")) %>% 
    mutate(log2_read_counts = log2(normalized_read_count)) %>%
    ggplot(aes(x = normalized_read_count, color = bin, fill = bin)) +
    geom_density(alpha = 0.1) + 
    labs(title = "Density Plot of Normalized Read Counts (STMN2 Screen)",
       x = "Normalized read counts (log2 scaled)",
       y = "Density") +
    theme_minimal()
stmn2_low_neg_ctrl_sgrna_counts <- neg_ctrl_stmn2_sgrna_sep %>%
  dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
  dplyr::filter(startsWith(bin, "low")) %>% 
    mutate(log2_read_counts = log2(normalized_read_count)) %>%
    ggplot(aes(x = normalized_read_count, color = bin, fill = bin)) +
    geom_density(alpha = 0.1) + 
    labs(title = "Density Plot of Normalized Read Counts (STMN2 Screen)",
       x = "Normalized read counts (log2 scaled)",
       y = "Density") +
    theme_minimal()
# Arrange plots side by side
gridExtra::grid.arrange(stmn2_high_neg_ctrl_sgrna_counts, stmn2_low_neg_ctrl_sgrna_counts, ncol = 2)

neg_ctrl_stmn2_sgrna_sep %>%
  dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
    mutate(log2_read_counts = log2(normalized_read_count)) %>%
    ggplot(aes(x = normalized_read_count, color = bin, fill = bin)) +
    geom_density(alpha = 0.1) + 
    labs(title = "Density Plot of Normalized Read Counts (STMN2 Screen)",
       x = "Normalized read counts (log2 scaled)",
       y = "Density") +
    theme_minimal()

# Group by 'bin' and calculate the median of 'normalized_read_count'
neg_ctrl_stmn2_sgrna_sep %>%
  group_by(bin) %>%
  dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
  dplyr::summarize(median_normalized_read_count = median(normalized_read_count, na.rm = TRUE))


# # Add a new column to categorize bins into "high" and "low"
# filtered_data <- neg_ctrl_stmn2_sgrna_sep %>%
#   dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
#   mutate(bin_category = ifelse(startsWith(bin, "high"), "high", "low"))

# TODO ----
  # plot hdr_mean col for neg ctrl sgRNAs from stmn2 screen - only generates 1 line though since it represents the geometric mean
  
  # compare plots of the hdr_mean cols of neg_ctrls from the stmn2 screen to the tdp screen
  # 

# tdp_sgrna_sep <- tdp_mageck_sgrna_df %>%
#   dplyr::select(sgrna, Gene, control_count, treatment_count) %>% 
#   separate(control_count, into = c("low1_rep1", "low2_rep2", "low3_rep3"), sep = "/") %>% 
#   separate(treatment_count, into = c("high1_rep1", "high2_rep2", "high3_rep3"), sep = "/") %>% 
#   pivot_longer(cols = starts_with(start_strings),
#                names_to = "bin",
#                values_to = "normalized_read_count")

  # TODO: consider different methods of normalization ----
stmn2_mageck_sgrna_df %>% 
  dplyr::filter(Gene=="negative_control")  # only select negative controls

```
# Try with signed log and cube root
```{r}
neg_ctrl_stmn2_sgrna_sep %>%
  dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
  mutate(signed_log_read_count = sign(normalized_read_count) * log(abs(normalized_read_count) + 1)) %>% 
    ggplot(aes(x = signed_log_read_count, color = bin, fill = bin)) +
    geom_density(alpha = 0.1) + 
    labs(title = "Density Plot of Signed Log Transformed Normalized Read Counts (STMN2 Screen)",
       x = "Signed Log Transformed Normalized Read Counts",
       y = "Density") +
    theme_minimal()

neg_ctrl_stmn2_sgrna_sep %>%
  dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
  mutate(cube_root_read_count = sign(normalized_read_count) * abs(normalized_read_count)^(1/3)) %>% 
    ggplot(aes(x = cube_root_read_count, color = bin, fill = bin)) +
    geom_density(alpha = 0.1) + 
    labs(title = "Density Plot of Cube Root Transformed Normalized Read Counts (STMN2 Screen)",
       x = "Cube Root Transformed Normalized Read Counts",
       y = "Density") +
    theme_minimal()


```


# Visualize Null Distribution of Non-targeting guides
```{r}
# Separate the normalized read count cols
neg_ctrl_tdp_sgrna <- tdp_mageck_sgrna_df %>%
  dplyr::select(sgrna, Gene, low = control_count, high = treatment_count) %>% 
  # tidyr::separate(control_count, into = c("low1_rep1", "low2_rep2", "low3_rep3"), sep = "/") %>% 
  # tidyr::separate(treatment_count, into = c("high1_rep1", "high2_rep2", "high3_rep3"), sep = "/") %>% 
  tidyr::pivot_longer(cols = starts_with(start_strings),
               names_to = "bin",
               values_to = "normalized_read_count") %>% 
  dplyr::filter(Gene=="negative_control")

# Create plots high/low FACS bins of normalized-log2 scaled sgRNA read counts
neg_ctrl_tdp_log2 <- neg_ctrl_tdp_sgrna %>%
  dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
    mutate(log2_read_counts = log2(normalized_read_count)) %>%
    ggplot(aes(x = normalized_read_count, color = bin, fill = bin)) +
    geom_density(alpha = 0.1) + 
    labs(title = "Density Plot of Normalized Read Counts (TDP Screen)",
       x = "Normalized read counts (log2 scaled)",
       y = "Density") +
    theme_minimal()



neg_ctrl_tdp_sgrna %>%
  dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
  mutate(signed_log_read_count = sign(normalized_read_count) * log(abs(normalized_read_count) + 1)) %>% 
    ggplot(aes(x = signed_log_read_count, color = bin, fill = bin)) +
    geom_density(alpha = 0.1) + 
    labs(title = "Density Plot of Signed Log Transformed Normalized Read Counts (TDP Screen)",
       x = "Signed Log Transformed Normalized Read Counts",
       y = "Density") +
    theme_minimal()

neg_ctrl_tdp_sgrna %>%
  dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
  mutate(cube_root_read_count = sign(normalized_read_count) * abs(normalized_read_count)^(1/3)) %>% 
    ggplot(aes(x = cube_root_read_count, color = bin, fill = bin)) +
    geom_density(alpha = 0.1) + 
    labs(title = "Density Plot of Cube Root Transformed Normalized Read Counts (TDP Screen)",
       x = "Cube Root Transformed Normalized Read Counts",
       y = "Density") +
    theme_minimal()

```

```{r Try replacing MAGeCK with Bayesian Modeling}
# Data Normalization ----
library(dplyr)
library(rstan)
# Example data normalization using Bayesian hierarchical modeling
# Assuming `data` is a dataframe with read counts

# Define the Bayesian model for normalization
stan_model <- "
data {
  int<lower=0> N; // number of observations
  int<lower=0> J; // number of groups (e.g., sgRNAs)
  int<lower=1,upper=J> group[N]; // group indicator
  real y[N]; // read counts
}
parameters {
  real<lower=0> sigma; // standard deviation
  real mu[J]; // group means
}
model {
  y ~ normal(mu[group], sigma);
}
"
stan_data <- list(N = nrow(data), J = length(unique(data$sgRNA)), group = as.numeric(factor(data$sgRNA)), y = data$read_count)
fit <- stan(model_code = stan_model, data = stan_data)

# Variance Estimation ----
# Extract posterior estimates
posterior <- extract(fit)
mu_estimates <- apply(posterior$mu, 2, mean)
sigma_estimate <- mean(posterior$sigma)

# Significance Testing ----
# Compute Bayesian p-values or credible intervals
# Assuming `test_data` and `control_data` are subsets of the original data
p_values <- sapply(1:length(mu_estimates), function(i) {
  mu_diff <- mu_estimates[i] - mean(control_data$read_count)
  p_value <- 2 * min(mean(mu_diff > 0), mean(mu_diff < 0))
  return(p_value)
})

# Rank Aggregation ----
# Rank genes by their posterior probabilities
gene_ranks <- data.frame(gene = unique(data$gene), p_value = p_values)
gene_ranks <- gene_ranks %>% arrange(p_value)

# FDR ----
# Adjust p-values for multiple testing using Bayesian FDR methods
library(qvalue)
q <- qvalue(p = gene_ranks$p_value)
gene_ranks <- gene_ranks %>% mutate(q_value = q$qvalues)

```

```{r}
# Load necessary libraries
library(dplyr)
library(rstan)

# Simulate example data
set.seed(123)
data <- tibble::tibble(
  sgRNA = rep(paste0("sgRNA_", 1:10), each = 6),
  gene = rep(paste0("Gene_", 1:10), each = 6),
  condition = rep(c("control", "treatment"), each = 3, times = 10),
  read_count = c(
    rpois(30, lambda = 100),  # control
    rpois(30, lambda = 120)   # treatment
  )
)

print(head(data))
# Define the Bayesian model for normalization
stan_model <- "
data {
  int<lower=0> N; // number of observations
  int<lower=0> J; // number of sgRNAs
  int<lower=1,upper=J> group[N]; // sgRNA indicator
  real y[N]; // read counts
}
parameters {
  real<lower=0> sigma; // standard deviation
  real mu[J]; // sgRNA means
}
model {
  y ~ normal(mu[group], sigma);
}
"

# Prepare data for Stan
stan_data <- list(
  N = nrow(data),
  J = length(unique(data$sgRNA)),
  group = as.numeric(factor(data$sgRNA)),
  y = log(data$read_count + 1)  # log-transform read counts
)

# Fit the model using Stan
fit <- stan(model_code = stan_model, data = stan_data, iter = 1000, chains = 4)

# Extract posterior estimates
posterior <- extract(fit)
mu_estimates <- apply(posterior$mu, 2, mean)
sigma_estimate <- mean(posterior$sigma)


# Compute differences between treatment and control
control_means <- mu_estimates[1:10]
treatment_means <- mu_estimates[11:20]

p_values <- sapply(1:10, function(i) {
  diff <- treatment_means[i] - control_means[i]
  p_value <- 2 * min(mean(diff > 0), mean(diff < 0))
  return(p_value)
})

# Create a data frame of results
results <- tibble::tibble(
  gene = unique(data$gene),
  control_mean = control_means,
  treatment_mean = treatment_means,
  p_value = p_values
)

print(results)
# Adjust p-values using the Benjamini-Hochberg method
results <- results %>%
  arrange(p_value) %>%
  mutate(q_value = p.adjust(p_value, method = "BH"))

print(results)

```

