---
title: "Dual-guide CRISPR Optimization"
output:
  html_document:
    df_print: paged
  html_notebook:
    fig_width: 8
    fig_height: 6.5
    fig_caption: true
---
# Setup
```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(qvalue)
library(corrplot)
library(gridExtra)
```

# Import MAGeCK summary files
```{r, echo=FALSE}
# Test best normalization method

# Import various screen data results ####
# Import STMN2 screen data
# 
sgrna_stmn2 <- read.delim("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/data_processed/rra-High-Low-STMN2.sgrna_summary.txt", sep = "\t", header = TRUE)
gene_stmn2 <- read.table("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/data_processed/rra-High-Low-STMN2.gene_summary.txt", sep = "\t", header = TRUE)

# Raw counts data
raw_counts <- read.delim("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/files_from_others/counts/all.count.txt", sep = "\t", header = TRUE)

# Import TDP-43 screen data
sgrna_tdp <- read.table("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/data_processed/tdp_i3N.sgrna_summary.txt", sep = "\t", header = TRUE)
# Import other screen data
sgrna_ipsc <- read.delim("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/data_processed/CP6277_rra-ipsc.sgrna_summary.txt", sep = "\t", header = TRUE)
sgrna_i3N <- read.delim("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/data_processed/CP6277_rra-i3N.sgrna_summary.txt", sep = "\t", header = TRUE)

# sgRNA hdr key: 
#   control_count: Normalized read counts in ctrl samples;
#   treatment_count: Normalized read counts in tx samples;
#   control_mean: Median read counts in ctrl samples;
#   treat_mean: Median read counts in tx samples

# Separate out all countrol_count and treatment_count columns as they may contain multiple values from replicate samples
# Create start_strings to select both high and low cols
start_strings <- c("high", "low")
# Separate the normalized read count cols
neg_ctrl_sgrna_stmn2 <- sgrna_stmn2 %>%
  dplyr::select(sgrna, Gene, control_count, treatment_count) %>% 
  tidyr::separate(control_count, into = c("low1_rep1", "low2_rep2", "low3_rep3"), sep = "/") %>% 
  tidyr::separate(treatment_count, into = c("high1_rep1", "high2_rep2", "high3_rep3"), sep = "/") %>% 
  tidyr::pivot_longer(cols = starts_with(start_strings),
                      names_to = "bin",
                      values_to = "normalized_read_count") %>% 
  dplyr::filter(Gene=="negative_control") %>% 
    mutate(normalized_read_count = as.numeric(normalized_read_count)) %>%
    mutate(log2_read_counts = log2(normalized_read_count),
           signed_log_read_count = sign(normalized_read_count) * log(abs(normalized_read_count) + 1),
           cube_root_read_count = sign(normalized_read_count) * abs(normalized_read_count)^(1/3))
# TDP dataframe
neg_ctrl_sgrna_tdp <- sgrna_tdp %>%
  dplyr::select(sgrna, Gene, low = control_count, high = treatment_count) %>% 
  tidyr::pivot_longer(cols = starts_with(start_strings),
                      names_to = "bin",
                      values_to = "normalized_read_count") %>% 
  dplyr::filter(Gene=="negative_control") %>% 
    mutate(normalized_read_count = as.numeric(normalized_read_count)) %>%
    mutate(log2_read_counts = log2(normalized_read_count),
           signed_log_read_count = sign(normalized_read_count) * log(abs(normalized_read_count) + 1),
           cube_root_read_count = sign(normalized_read_count) * abs(normalized_read_count)^(1/3))
# i3N dataframe
neg_ctrl_sgrna_i3N <- sgrna_i3N %>%
  dplyr::select(sgrna, Gene, control_count, treatment_count) %>% 
  tidyr::separate(control_count, into = c("low1_rep1", "low2_rep2"), sep = "/") %>% 
  tidyr::separate(treatment_count, into = c("high1_rep1", "high2_rep2"), sep = "/") %>% 
  tidyr::pivot_longer(cols = starts_with(start_strings),
                      names_to = "bin",
                      values_to = "normalized_read_count") %>% 
  dplyr::filter(Gene=="negative_control") %>% 
    mutate(normalized_read_count = as.numeric(normalized_read_count)) %>%
    mutate(log2_read_counts = log2(normalized_read_count),
           signed_log_read_count = sign(normalized_read_count) * log(abs(normalized_read_count) + 1),
           cube_root_read_count = sign(normalized_read_count) * abs(normalized_read_count)^(1/3))
# iPSC dataframe
neg_ctrl_sgrna_ipsc <- sgrna_ipsc %>%
  dplyr::select(sgrna, Gene, control_count, treatment_count) %>% 
  tidyr::separate(control_count, into = c("low1_rep1", "low2_rep2"), sep = "/") %>% 
  tidyr::separate(treatment_count, into = c("high1_rep1", "high2_rep2"), sep = "/") %>% 
  tidyr::pivot_longer(cols = starts_with(start_strings),
                      names_to = "bin",
                      values_to = "normalized_read_count") %>% 
  dplyr::filter(Gene=="negative_control") %>% 
    mutate(normalized_read_count = as.numeric(normalized_read_count)) %>%
    mutate(log2_read_counts = log2(normalized_read_count),
           signed_log_read_count = sign(normalized_read_count) * log(abs(normalized_read_count) + 1),
           cube_root_read_count = sign(normalized_read_count) * abs(normalized_read_count)^(1/3))
```

# Dealing with Raw Read Counts
```{r Transform raw counts, echo=FALSE}
# TODO: translate count norm functions into R
start_strings = c('Low', 'High')

# Raw counts
neg_ctrl_raw_counts_long <- raw_counts %>%
  tidyr::pivot_longer(cols = starts_with(start_strings),
                      names_to = "bin",
                      values_to = "normalized_read_count") %>% 
  dplyr::filter(Gene=="negative_control") %>%
    mutate(log2_read_counts = log2(normalized_read_count),
           signed_log_read_count = sign(normalized_read_count) * log(abs(normalized_read_count) + 1),
           cube_root_read_count = sign(normalized_read_count) * abs(normalized_read_count)^(1/3))

neg_ctrl_raw_counts_wide <- raw_counts %>%
  dplyr::filter(Gene=="negative_control")

```

# Try recreating MAGeCK CountNorm function
```{r CountNorm setup, echo=FALSE}
# Equivalent to NormMsgs class in Python
NormMsgs <- function() {
  list(haswarning = FALSE)
}

mageckcount_gettotalnormfactor <- function(ctable) {
  # Number of samples
  n <- length(ctable[[1]])
  # Number of sgRNAs
  m <- length(ctable)
  
  # Calculate the sum for each sample
  sumsample <- rep(0, n)
  for (v in ctable) {
    sumsample <- sumsample + unlist(v)
  }
  
  # Calculate the average sample sum and normalization factors
  avgsample <- sum(sumsample) / n
  samplefactor <- avgsample / sumsample
  
  return(samplefactor)
}

mageckcount_getzerocounts <- function(ctable) {
  # Number of samples
  n <- length(ctable[[1]])
  # Number of sgRNAs
  m <- length(ctable)
  
  zerofactor <- rep(0, n)
  for (ni in 1:n) {
    nzr <- sapply(ctable, function(v) if (v[ni] == 0) 1 else 0)
    zerofactor[ni] <- sum(nzr)
  }
  
  return(zerofactor)
}

# mageckcount_getmediannormfactor <- function(ctable) {
#   # Number of samples
#   n <- length(ctable[[1]])
#   # Number of sgRNAs
#   m <- length(ctable)
#   
#   # Calculate the geometric mean for each sgRNA
#   meanval <- sapply(ctable, function(v) {
#     if (sum(v) > 0) {
#       exp(sum(log(v + 1)) / n)
#     } else {
#       NA
#     }
#   })
#   
#   # Replace non-positive geometric means with 1
#   meanval <- sapply(meanval, function(v) ifelse(is.na(v) || v <= 0, 1, v))
#   
#   medianfactor <- rep(0, n)
#   for (ni in 1:n) {
#     meanfactor <- sapply(ctable, function(v) v[ni] / meanval[names(ctable) == names(v)])
#     xfactor <- median(meanfactor, na.rm = TRUE)
#     
#     if (xfactor > 0.0) {
#       medianfactor[ni] <- 1.0 / xfactor
#     }
#   }
#   
#   return(medianfactor)
# }

mageckcount_getmediannormfactor <- function(ctable) {
  # Number of samples
  n <- length(ctable[[1]])
  # Number of sgRNAs
  m <- length(ctable)
  
  # Calculate the geometric mean for each sgRNA
  meanval <- sapply(ctable, function(v) {
    if (sum(v) > 0) {
      exp(sum(log(v + 1)) / n)
    } else {
      NA
    }
  })
  
  # Replace non-positive geometric means with 1
  meanval <- sapply(meanval, function(v) ifelse(is.na(v) || v <= 0, 1, v))
  
  medianfactor <- rep(0, n)
  for (ni in 1:n) {
    meanfactor <- sapply(ctable, function(v) v[ni] / meanval[names(ctable) == names(v)])
    
    # Ensure that the meanfactor is unlisted to avoid the atomic error
    meanfactor <- unlist(meanfactor)
    
    # Remove NA values from meanfactor before sorting
    meanfactor <- meanfactor[!is.na(meanfactor)]
    
    # Sort and get the median value only if meanfactor has elements
    if (length(meanfactor) > 0) {
      xfactor <- sort(meanfactor)[ceiling(length(meanfactor) / 2)]
    } else {
      xfactor <- 0
    }
    
    if (xfactor > 0.0) {
      medianfactor[ni] <- 1.0 / xfactor
    }
  }
  
  return(medianfactor)
}


normalizeCounts <- function(ctable, sgdict = NULL, method = 'median', returnfactor = FALSE,  reversefactor = FALSE, controlsgfile = NULL, norm_msgs = NULL) {
  
  # Initialize norm_msgs if not provided
  if (is.null(norm_msgs)) {
    norm_msgs <- list(haswarning = FALSE)
  }
  
  # If ctable is empty, return a copy
  if (length(ctable) == 0) {
    return(ctable)
  }
  
  # Calculate the number of samples and sgRNAs
  n <- length(ctable[[1]])  # Number of samples
  m <- length(ctable)  # Number of sgRNAs
  
  # Handle control normalization
  if (method == 'control') {
    if (is.null(controlsgfile)) {
      stop("Error: a list of control sgRNAs should be specified when using control normalization.")
    }
    
    controlsglist <- readLines(controlsgfile)
    message("Loaded ", length(controlsglist), " control sgRNAs from ", controlsgfile)
    
    ctable_nm <- ctable[controlsglist]
    
    if (length(ctable_nm) == 0 && !is.null(sgdict)) {
      ctable_nm <- ctable[sapply(names(ctable), function(k) sgdict[[k]][1] %in% controlsglist)]
    }
    
    message("control sgRNAs for normalization: ", length(ctable_nm))
    
    if (length(ctable_nm) == 0) {
      stop("Error: cannot find records of control sgRNAs in the read count table.")
    }
    
    method <- 'median'
  } else {
    ctable_nm <- ctable
  }
  
  # Calculate the total normalization factor
  samplefactor <- mageckcount_gettotalnormfactor(ctable_nm)
  message("Initial (total) size factor: ", paste(samplefactor, collapse = " "))
  
  nzeros <- mageckcount_getzerocounts(ctable_nm)
  
  if (method == 'median') {
    medianfactor <- mageckcount_getmediannormfactor(ctable_nm)
    usetotalnorm <- FALSE
    
    for (ni in 1:n) {
      if (medianfactor[ni] == 0.0) {
        warning("Sample ", ni, " has zero median count, so median normalization is not possible. Switching to total read count normalization.")
        norm_msgs$haswarning <- TRUE
        usetotalnorm <- TRUE
      }
      
      zerof <- nzeros[ni] / length(ctable_nm)
      if (zerof > 0.45) {
        warning("Sample ", ni, " has too many zero-count sgRNAs (", zerof, "), and median normalization is unstable. Switching to total read count normalization.")
        norm_msgs$haswarning <- TRUE
        usetotalnorm <- TRUE
      } else if (zerof > 0.3) {
        warning("Sample ", ni, " has too many zero-count sgRNAs (", zerof, "), and median normalization may be unstable. Consider switching to total read count normalization.")
        norm_msgs$haswarning <- TRUE
      }
    }
    
    if (!usetotalnorm) {
      samplefactor <- medianfactor
      message("Median factor: ", paste(samplefactor, collapse = " "))
    }
  } else if (method == 'none') {
    samplefactor <- rep(1.0, n)
  }
  
  message("Final size factor: ", paste(samplefactor, collapse = " "))
  
  warn_norm <- any(samplefactor < 0.2 | samplefactor > 5)
  if (warn_norm) {
    warning("Some samples have unusually small or large normalization factors (<0.2 or >5). Please double-check the normalization process.")
    norm_msgs$haswarning <- TRUE
  }
  
  if (returnfactor) {
    if (reversefactor) {
      return(1.0 / samplefactor)
    } else {
      return(samplefactor)
    }
  }
  
  # Normalize the table
  ntable <- lapply(ctable, function(v) samplefactor * v)
  return(ntable)
}


```

## Try creating new bayesian inference framework
```{r bayes, echo=FALSE, include=FALSE}
# Low will be control grp
C_counts <- raw_counts %>% 
  dplyr::select(Low_1, Low_2, Low_3)

# High will be treatment grp
T_counts <- raw_counts %>% 
  dplyr::select(High_1, High_2, High_3)

calculate_lfc <- function(T_counts, C_counts, df) {
  # T_counts and C_counts should be matrices where each row corresponds to a guide 
  # and each column corresponds to a replicate. Rows correspond to different guides.
  # Columns correspond to replicates (RT for treatment, RC for control).
  
  # Calculate Tg,i,l,r and Cg,i,r for each guide and replicate. Drop first two columns as they contain gene and sgRNA info
  T_log <- log2(T_counts + 32)
  C_log <- log2(C_counts + 32)
  
  # Median normalization across guides for each replicate
  T_median <- apply(T_log, 2, median)
  C_median <- apply(C_log, 2, median)
  
  # Subtract the median for each replicate
  T_median_normalized <- sweep(T_log, 2, T_median, "-")
  C_median_normalized <- sweep(C_log, 2, C_median, "-")
  
  # Calculate the average across replicates for both treatment and control
  T_mean <- rowMeans(T_median_normalized)
  C_mean <- rowMeans(C_median_normalized)
  
  # Calculate the observed log2 fold change
  y <- T_mean - C_mean

  # Add the calculated vectors to the dataframe as new columns
  df$T_log <- T_log
  df$C_log <- C_log
  df$T_median <- T_median  # Average of the medians across treatment replicates
  df$C_median <- C_median  # Average of the medians across control replicates
  df$T_median_normalized <- T_median_normalized
  df$C_median_normalized <- C_median_normalized
  df$T_mean <- T_mean
  df$C_mean <- C_mean
  # df$T_mean_normalized <- rowMeans(T_median_normalized)   # Treatment mean of median normalized
  # df$C_mean_normalized <- rowMeans(C_median_normalized)   # Control mean of median normalized
  df$LFC <- y                        # Log2 fold change
  
  return(df)
}

# Function to calculate priors for the Gaussian distribution
calculate_priors <- function(y_gil, N_g, k = 0.5) {
  # Initialize prior parameters
  mu_w <- 0
  sigma_w2 <- 1000
  mu_x <- 0
  sigma_x2 <- 1
  
  # Calculate the smoothed variance estimates
  T_var <- apply(T_counts, 1, var)
  C_var <- apply(C_counts, 1, var)
  
  # Prior precision based on the sum of treatment and control variances
  tau_gil <- (T_var + C_var)
  
  # Assign priors
  ag <- k * T_var
  bg <- k * (T_var + C_var)
  
  return(list(mu_w = mu_w, sigma_w2 = sigma_w2, mu_x = mu_x, sigma_x2 = sigma_x2, tau_gil = tau_gil, ag = ag, bg = bg))
}


install.packages("MASS")
install.packages("scipy")
install.packages("gaussianKde")

set.seed(42)  # For reproducibility

select_random_gene <- function(gene_library) {
  return(sample(names(gene_library), 1))
}

select_gRNAs <- function(gene_library, gene, N) {
  return(sample(gene_library[[gene]], N, replace = TRUE))
}
select_negative_controls <- function(negative_control_set, N) {
  return(sample(negative_control_set, N, replace = TRUE))
}
generate_negative_pseudogenes <- function(gene_library, negative_control_set, N, num_pseudogenes = 2000) {
  negative_pseudogenes <- vector("list", num_pseudogenes)
  
  for (i in 1:num_pseudogenes) {
    random_gene <- select_random_gene(gene_library)
    selected_gRNAs <- select_gRNAs(gene_library, random_gene, N)
    negative_pseudogenes[[i]] <- select_negative_controls(negative_control_set, N)
  }
  
  return(negative_pseudogenes)
}


# Function to generate negative pseudogenes by resampling
generate_pseudogenes <- function(gRNA_count, replicates, n_pseudogenes = 2000) {
  pseudo_results <- matrix(0, nrow = n_pseudogenes, ncol = replicates)
  
  for (i in 1:n_pseudogenes) {
    pseudo_results[i, ] <- sample(rnorm(gRNA_count, mean = 0, sd = 1), replicates)
  }
  
  return(pseudo_results)
}















# Calculate LFC
y <- calculate_lfc(T_counts, C_counts, raw_counts)


priors <- calculate_priors(y, N_g = 10)
print(priors)

# Example resampling
pseudo_data <- generate_pseudogenes(gRNA_count = 10, replicates = 5)
print(pseudo_data)



```


## Test R-adapted CountNorm function
### TODO: test with raw sgRNA counts
```{r test RCountNorm, echo=FALSE}
# Sample data: 5 sgRNAs (rows) and 3 samples (columns)
ctable <- list(
  sgRNA1 = c(100, 200, 300),
  sgRNA2 = c(400, 500, 600),
  sgRNA3 = c(0, 0, 0),  # This sgRNA has zero counts
  sgRNA4 = c(800, 700, 600),
  sgRNA5 = c(1000, 900, 800)
)

# Define a dummy sgdict (not used in this example)
sgdict <- list(
  sgRNA1 = c("sgRNA1_ID", "Gene1"),
  sgRNA2 = c("sgRNA2_ID", "Gene2"),
  sgRNA3 = c("sgRNA3_ID", "Gene3"),
  sgRNA4 = c("sgRNA4_ID", "Gene4"),
  sgRNA5 = c("sgRNA5_ID", "Gene5")
)

# Normalization method: median
method <- 'median'

# Call the normalizeCounts function
normalized_ctable <- normalizeCounts(ctable, sgdict = sgdict, method = method)

# Print the original and normalized count tables
print("Original Counts:")
print(ctable)

print("Normalized Counts:")
print(normalized_ctable)

```

# Visualize non-targeting guides
```{r Visualize best normalization method, fig.width=10, fig.height=8,  echo=FALSE, warning=FALSE}
# Visualize Null Distribution of Non-targeting guides ####
# Function to plot the log2 transformed values
log2_density_plot <- function(data_frame, title_label) {
  data_frame %>%
    dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
    mutate(log2_read_counts = log2(normalized_read_count)) %>%
    ggplot(aes(x = normalized_read_count, color = bin, fill = bin)) +
    geom_density(alpha = 0.1) + 
    labs(title = title_label,
         x = "Log2 Transformed \n Normalized Read Counts",
         y = "Density") +
    theme_minimal()
}
# Function to plot the signed log transformation
signed_log_density_plot <- function(data_frame, title_label) {
  data_frame %>%
    mutate(normalized_read_count = as.numeric(normalized_read_count)) %>%
    mutate(signed_log_read_count = sign(normalized_read_count) * log(abs(normalized_read_count) + 1)) %>% 
    ggplot(aes(x = signed_log_read_count, color = bin, fill = bin)) +
    geom_density(alpha = 0.1) + 
    labs(title = title_label,
         x = "Signed Log Transformed \n Normalized Read Counts",
         y = "Density") +
    theme_minimal()
}
# Function to plot the cube root transformation
cube_root_density_plot <- function(data_frame, title_label) {
  data_frame %>%
    mutate(normalized_read_count = as.numeric(normalized_read_count)) %>%
    mutate(cube_root_read_count = sign(normalized_read_count) * abs(normalized_read_count)^(1/3)) %>%
    ggplot(aes(x = cube_root_read_count, color = bin, fill = bin)) +
    geom_density(alpha = 0.1) + 
    labs(title = title_label,
         x = "Cube Root Transformed \n Normalized Read Counts",
         y = "Density") +
    theme_minimal()
}

stmn2_log2_density_plot <- log2_density_plot(neg_ctrl_sgrna_stmn2, "STMN2 Screen")
tdp_log2_density_plot <- log2_density_plot(neg_ctrl_sgrna_tdp, "TDP Screen")
iN3_log2_density_plot <- log2_density_plot(neg_ctrl_sgrna_i3N, "i3N Screen")
ipsc_log2_density_plot <- log2_density_plot(neg_ctrl_sgrna_ipsc, "iPSC Screen")

stmn2_cuberoot_density_plot <- cube_root_density_plot(neg_ctrl_sgrna_stmn2, "STMN2 Screen")
tdp_cuberoot_density_plot <- cube_root_density_plot(neg_ctrl_sgrna_tdp, "TDP Screen")
iN3_cuberoot_density_plot <- cube_root_density_plot(neg_ctrl_sgrna_i3N, "i3N Screen")
ipsc_cuberoot_density_plot <- cube_root_density_plot(neg_ctrl_sgrna_ipsc, "iPSC Screen")

stmn2_signedlog_density_plot <- signed_log_density_plot(neg_ctrl_sgrna_stmn2, "STMN2 Screen")
tdp_signedlog_density_plot <- signed_log_density_plot(neg_ctrl_sgrna_tdp, "TDP Screen")
iN3_signedlog_density_plot <- signed_log_density_plot(neg_ctrl_sgrna_i3N, "i3N Screen")
ipsc_signedlog_density_plot <- signed_log_density_plot(neg_ctrl_sgrna_ipsc, "iPSC Screen")

stmn2_log2_raw_density_plot <- log2_density_plot(neg_ctrl_raw_counts, "STMN2 Raw Counts")

stmn2_cuberoot_raw_density_plot <- cube_root_density_plot(neg_ctrl_raw_counts, "STMN2 Raw Counts")

stmn2_signedlog_raw_density_plot <- signed_log_density_plot(neg_ctrl_raw_counts, "STMN2 Raw Counts")


# Arrange plots side by side
gridExtra::grid.arrange(stmn2_log2_raw_density_plot, stmn2_log2_density_plot, tdp_log2_density_plot, iN3_log2_density_plot, ipsc_log2_density_plot, ncol = 3, nrow = 2)
gridExtra::grid.arrange(stmn2_cuberoot_raw_density_plot, stmn2_cuberoot_density_plot, tdp_cuberoot_density_plot, iN3_cuberoot_density_plot, ipsc_cuberoot_density_plot, ncol = 3, nrow = 2)
gridExtra::grid.arrange(stmn2_signedlog_raw_density_plot, stmn2_signedlog_density_plot, tdp_signedlog_density_plot, iN3_signedlog_density_plot, ipsc_signedlog_density_plot, ncol = 3, nrow = 2)

# Alternate arrangement by screen
gridExtra::grid.arrange(stmn2_log2_raw_density_plot, stmn2_cuberoot_raw_density_plot, stmn2_signedlog_raw_density_plot, stmn2_log2_density_plot, stmn2_cuberoot_density_plot, stmn2_signedlog_density_plot, ncol = 3, nrow = 2)
gridExtra::grid.arrange(, ncol = 3, nrow = 1)
gridExtra::grid.arrange(tdp_log2_density_plot, tdp_cuberoot_density_plot, tdp_signedlog_density_plot, ncol = 3, nrow = 2)
gridExtra::grid.arrange(iN3_log2_density_plot, iN3_cuberoot_density_plot, iN3_signedlog_density_plot, ncol = 3, nrow = 1)
gridExtra::grid.arrange(ipsc_log2_density_plot, ipsc_cuberoot_density_plot, ipsc_signedlog_density_plot, ncol = 3, nrow = 1)

library(patchwork)
combined_stmn2 <- stmn2_log2_raw_density_plot+ stmn2_cuberoot_raw_density_plot+ stmn2_signedlog_raw_density_plot+ stmn2_log2_density_plot+ stmn2_cuberoot_density_plot+ stmn2_signedlog_density_plot & theme(legend.position = "right")
combined + plot_layout(guides = "collect")

combined_ipsc <- iN3_log2_density_plot+ iN3_cuberoot_density_plot+ iN3_signedlog_density_plot+ ipsc_log2_density_plot+ ipsc_cuberoot_density_plot+ ipsc_signedlog_density_plot & theme(legend.position = "right")
combined_ipsc + plot_layout(guides = "collect")

```

```{r scratch Alternate visualize best normalization method, echo=FALSE, eval=FALSE, include=FALSE, warning=FALSE}

# Create plots high/low FACS bins of normalized-log2 scaled sgRNA read counts
# neg_ctrl_sgrna_stmn2_high <- neg_ctrl_sgrna_stmn2 %>%
#   dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
#   dplyr::filter(startsWith(bin, "high")) %>% 
#   mutate(log2_read_counts = log2(normalized_read_count)) %>%
#   ggplot(aes(x = normalized_read_count, color = bin, fill = bin)) +
#   geom_density(alpha = 0.1) + 
#   labs(title = "Density Plot of Normalized Read Counts (STMN2 Screen)",
#        x = "Normalized read counts (log2 scaled)",
#        y = "Density") +
#   theme_minimal()
# neg_ctrl_sgrna_stmn2_low <- neg_ctrl_sgrna_stmn2 %>%
#   dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
#   dplyr::filter(startsWith(bin, "low")) %>% 
#   mutate(log2_read_counts = log2(normalized_read_count)) %>%
#   ggplot(aes(x = normalized_read_count, color = bin, fill = bin)) +
#   geom_density(alpha = 0.1) + 
#   labs(title = "Density Plot of Normalized Read Counts (STMN2 Screen)",
#        x = "Normalized read counts (log2 scaled)",
#        y = "Density") +
#   theme_minimal()

# Both high and low bins overlayed
# neg_ctrl_sgrna_stmn2 %>%
# dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
#   mutate(log2_read_counts = log2(normalized_read_count)) %>%
#   ggplot(aes(x = normalized_read_count, color = bin, fill = bin)) +
#   geom_density(alpha = 0.1) + 
#   labs(title = "Density Plot of Normalized Read Counts (STMN2 Screen)",
#        x = "Normalized read counts (log2 scaled)",
#        y = "Density") +
#   theme_minimal()

# Group by 'bin' and calculate the median of 'normalized_read_count'
# neg_ctrl_sgrna_stmn2 %>%
#   group_by(bin) %>%
#   dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
#   dplyr::summarize(median_normalized_read_count = median(normalized_read_count, na.rm = TRUE))

# TODO: consider different methods of normalization ----
# sgrna_stmn2 %>% 
#   dplyr::filter(Gene=="negative_control")  # only select negative controls

# Try with signed log and cube root
# neg_ctrl_sgrna_stmn2 %>%
#   dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
#   mutate(signed_log_read_count = sign(normalized_read_count) * log(abs(normalized_read_count) + 1)) %>% 
#   ggplot(aes(x = signed_log_read_count, color = bin, fill = bin)) +
#   geom_density(alpha = 0.1) + 
#   labs(title = "Density Plot of Signed Log Transformed Normalized Read Counts (STMN2 Screen)",
#        x = "Signed Log Transformed Normalized Read Counts",
#        y = "Density") +
#   theme_minimal()
# 
# neg_ctrl_sgrna_stmn2 %>%
#   dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
#   mutate(cube_root_read_count = sign(normalized_read_count) * abs(normalized_read_count)^(1/3)) %>% 
#   ggplot(aes(x = cube_root_read_count, color = bin, fill = bin)) +
#   geom_density(alpha = 0.1) + 
#   labs(title = "Density Plot of Cube Root Transformed Normalized Read Counts (STMN2 Screen)",
#        x = "Cube Root Transformed Normalized Read Counts",
#        y = "Density") +
#   theme_minimal()


# # Create plots high/low FACS bins of normalized-log2 scaled sgRNA read counts
# neg_ctrl_tdp_log2 <- neg_ctrl_tdp_sgrna %>%
#   dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
#   mutate(log2_read_counts = log2(normalized_read_count)) %>%
#   ggplot(aes(x = normalized_read_count, color = bin, fill = bin)) +
#   geom_density(alpha = 0.1) + 
#   labs(title = "Density Plot of Normalized Read Counts (TDP Screen)",
#        x = "Normalized read counts (log2 scaled)",
#        y = "Density") +
#   theme_minimal()
```


# Correlation matrix plot of non-targeting control guides
```{r, echo=FALSE, fig.width=10, fig.height=8}
# Create function for calculating correlations and outputting lower triangle correlation matrices ####
lower.tri.cor.matrix <- function(df){
  cor.matrix<-round(cor(df, use = "complete.obs"), 3)
  cor.matrix[upper.tri(cor.matrix)]<-""
  cor.matrix<-as.data.frame(cor.matrix)
  cor.matrix
  
  return(cor.matrix)
}
#####     CV VALUES    #####
# inter.assay.cv <- function(df, ...){
#   grp_vars <- quos(...)
#   selected.cols <- df %>%
#     select(!!!grp_vars)
#   dat.as.matrix <- as.matrix(selected.cols)
#   selected.cols$mean <- rowMeans(selected.cols,na.rm = T)
#   selected.cols$sd <- rowSds(dat.as.matrix,na.rm = T)
#   selected.cols <- selected.cols %>%
#     mutate(CV = sd/mean,
#            pct.CV = (sd/mean)*100)
#   
#   return(selected.cols)
# }
# R1.cv <- inter.assay.cv(joined_df, low1_rep1_stmn2, low1_rep1_i3N)
# names(R1.cv) <- c("R1_S1", "R1_S2", "R1_mean", "R1_sd", "R1_CV", "R1_pct.CV")
# mean(R1.cv$R1_pct.CV, na.rm = T)

inter.assay.cv <- function(df, ...) {
  grp_vars <- quos(...)
  selected.cols <- df %>%
    select(!!!grp_vars)
  
  dat.as.matrix <- as.matrix(selected.cols)
  selected.cols <- selected.cols %>%
    mutate(
      mean = rowMeans(selected.cols, na.rm = TRUE),
      sd = rowSds(dat.as.matrix, na.rm = TRUE),
      CV = sd / mean,
      pct.CV = (sd / mean) * 100
    )
  
  return(selected.cols)
}

neg_ctrl_sgrna_stmn2 <- sgrna_stmn2 %>%
  dplyr::select(sgrna, Gene, control_count, treatment_count) %>% 
  tidyr::separate(control_count, into = c("low1_rep1", "low2_rep2", "low3_rep3"), sep = "/") %>% 
  tidyr::separate(treatment_count, into = c("high1_rep1", "high2_rep2", "high3_rep3"), sep = "/") %>% 
  dplyr::filter(Gene=="negative_control")
colnames(neg_ctrl_sgrna_stmn2)[-(1:2)]<-paste(colnames(neg_ctrl_sgrna_stmn2)[-(1:2)],"stmn2",sep="_")
# TDP dataframe
neg_ctrl_sgrna_tdp <- sgrna_tdp %>%
  dplyr::select(sgrna, Gene, low = control_count, high = treatment_count) %>% 
  dplyr::filter(Gene=="negative_control")
colnames(neg_ctrl_sgrna_tdp)[-(1:2)]<-paste(names(neg_ctrl_sgrna_tdp)[-(1:2)],"tdp",sep="_")
# i3N dataframe
neg_ctrl_sgrna_i3N <- sgrna_i3N %>%
  dplyr::select(sgrna, Gene, control_count, treatment_count) %>% 
  tidyr::separate(control_count, into = c("low1_rep1", "low2_rep2"), sep = "/") %>% 
  tidyr::separate(treatment_count, into = c("high1_rep1", "high2_rep2"), sep = "/") %>% 
  dplyr::filter(Gene=="negative_control")
colnames(neg_ctrl_sgrna_i3N)[-(1:2)]<-paste(names(neg_ctrl_sgrna_i3N)[-(1:2)],"i3N",sep="_")
# iPSC dataframe
neg_ctrl_sgrna_ipsc <- sgrna_ipsc %>%
  dplyr::select(sgrna, Gene, control_count, treatment_count) %>% 
  tidyr::separate(control_count, into = c("low1_rep1", "low2_rep2"), sep = "/") %>% 
  tidyr::separate(treatment_count, into = c("high1_rep1", "high2_rep2"), sep = "/") %>% 
  dplyr::filter(Gene=="negative_control")
colnames(neg_ctrl_sgrna_ipsc)[-(1:2)]<-paste(names(neg_ctrl_sgrna_ipsc)[-(1:2)],"ipsc",sep="_")

# Put all negative control read count values into a single df
joined_df <- dplyr::left_join(neg_ctrl_sgrna_stmn2, neg_ctrl_sgrna_tdp)
joined_df <- dplyr::left_join(joined_df, neg_ctrl_sgrna_i3N)
joined_df <- dplyr::left_join(joined_df, neg_ctrl_sgrna_ipsc)

# Convert specified columns to numeric
joined_df <- joined_df %>%
 mutate(across(-c(1, 2), as.numeric))

neg_ctrl_cor <- cor(joined_df[,-c(1,2)])
# head(round(z, 2))
# Matrix of p-values with corresponding correlation matrix
# mat : is a matrix of data
# ... : further arguments to pass to the native R cor.test function
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
# matrix of the p-value of the correlation
p.mat <- cor.mtest(joined_df[,-c(1,2)])
col<- colorRampPalette(c("maroon1", "lightgreen"))(5)
# Save the original margins
original_margins <- par()$mar

corrplot(neg_ctrl_cor, method = "number", type = "lower", tl.col = "black", tl.srt = 10, tl.cex = 0.85, p.mat = p.mat, sig.level = 0.01, insig = "blank", col = col, is.corr = FALSE, title = "Correlation matrix of sgRNA guide counts", mar = c(0.1, 0, 1.5, 0) + 0.2)
# Restore the original margins
corrplot(neg_ctrl_cor, type = "lower", tl.col = "black", tl.srt = 10, tl.cex = 0.85, p.mat = p.mat, sig.level = 0.01, insig = "blank", col = col, is.corr = FALSE, title = "Correlation matrix of sgRNA guide counts", mar = c(0.1, 0.1, 1.5, 0.1) + 0.2)
corrplot(neg_ctrl_cor, method = "number", order = "hclust", type = "lower", tl.col = "black", tl.srt = 10,tl.cex = 0.85, p.mat = p.mat, sig.level = 0.01, insig = "blank", col = col, is.corr = FALSE, title = "Correlation matrix of sgRNA guide counts - hierarchical cluster", mar = c(0.1, 0.1, 1.5, 0.1) + 0.2)
corrplot(neg_ctrl_cor, type = "lower", order = "hclust", tl.col = "black", tl.srt = 10, tl.cex = 0.85, p.mat = p.mat, sig.level = 0.01, insig = "blank", col = col, is.corr = FALSE,  title = "Correlation matrix of sgRNA guide counts - hierarchical cluster", mar = c(0.1, 0.1, 1.5, 0.1) + 0.2)
```

```{r comparing BA plots.rmd, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# If this is your first time using R, please install packages with the commented lines below. Note the line will need to be uncommented (delete leading #) in order to run
# install.packages(c("ggplot2", "BlandAltmanLeh", "readxl", "stargazer", "matrixStats"), dependencies = TRUE)
# Load packages into the current R session.
library(ggplot2)  #data visualization
library(BlandAltmanLeh)
library(readxl)   #read excel files
library(stargazer)  #print out clean tables
library(matrixStats)
library(corrplot)
library(gridExtra)
library(dplyr)
library(ggExtra)
library(blandr)
library(ggpubr)
#####     CV VALUES    #####
inter.assay.cv <- function(df, ...){
  grp_vars <- quos(...)
  selected.cols <- df %>%
    select(!!!grp_vars)
  dat.as.matrix <- as.matrix(selected.cols)
  selected.cols$mean <- rowMeans(selected.cols,na.rm = T)
  selected.cols$sd <- rowSds(dat.as.matrix,na.rm = T)
  selected.cols <- selected.cols %>%
    mutate(CV = sd/mean,
           pct.CV = (sd/mean)*100)
  
  return(selected.cols)
}

```

# Calculate raw count differences between the bins of the STMN2 screen 
Create columns to store raw count differences between the STMN2 bins. Use to examine raw inTRA-assay variability
```{r raw count difference-intRA-assay variability, echo=FALSE, include=FALSE}
# STMN2
# Extract column names starting from the 3rd column
columns <- colnames(raw_counts)[3:ncol(raw_counts)]

# Create a list of all possible combinations of column comparisons
  # Generate all possible combinations of columns taken 2 at a time
comparisons <- combn(columns, 2, simplify = FALSE)

# Store the combinations as a list of lists and assign to the variable `samples`
samples <- lapply(comparisons, function(x) c(x[1], x[2]))

# Function to calculate differences between columns in df
calculate_differences <- function(pair) {
  col1 <- neg_ctrl_raw_counts_wide[[pair[[1]]]]
  col2 <- neg_ctrl_raw_counts_wide[[pair[[2]]]]
  difference <- col1 - col2
  return(difference)
}

# Apply the function to each pair in samples
differences_list <- lapply(samples, calculate_differences)

# Naming the list elements to reflect the column comparisons
names(differences_list) <- sapply(samples, function(pair) paste(pair[[1]], "-", pair[[2]]))

```

<!-- # Compare Raw sgRNA counts against counts in MAGeCK sgRNA summary file -->
```{r, echo=FALSE}
# sgrna_stmn2 <- read.delim("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/data_processed/rra-High-Low-STMN2.sgrna_summary.txt", sep = "\t", header = TRUE)
# # Import Raw counts file
# raw_counts_sgrna_stmn2 <- read.delim("/Users/Claire/Library/CloudStorage/Box-Box/MSTP Research Rotations/Summer_2024_Might_Ward_Lab/CRISPRi_screenfiles/CRISPRi_alg_optimization/counts/all.count.txt", sep = "\t", header = TRUE)
# 
# counts_sgrna_stmn2 <- sgrna_stmn2 %>%
#   dplyr::select(sgrna, Gene, control_count, treatment_count) %>% 
#   tidyr::separate(control_count, into = c("low1", "low2", "low3"), sep = "/") %>% 
#   tidyr::separate(treatment_count, into = c("high1", "high2", "high3"), sep = "/")
#   # tidyr::pivot_longer(cols = starts_with(start_strings),
#   #              names_to = "bin",
#   #              values_to = "normalized_read_count")
# 
# start_strings <- c("High", "Low")
# 
# raw_counts_sgrna_stmn2_long <- raw_counts_sgrna_stmn2 %>% 
#   tidyr::pivot_longer(cols = starts_with(start_strings),
#                       names_to = "bin",
#                       values_to = "raw_counts")
# raw_counts_sgrna_stmn2_long %>% 
#     ggplot(aes(x = raw_counts, color = bin, fill = bin)) +
#     geom_density(alpha = 0.1) + 
#     labs(x = "Raw Read Counts",
#          y = "Density") +
#     theme_minimal()
# 
# 
# raw_counts_sgrna_stmn2_long %>%
#     dplyr::mutate(normalized_read_count = as.numeric(normalized_read_count)) %>% 
#     mutate(log2_read_counts = log2(normalized_read_count)) %>%
#     ggplot(aes(x = normalized_read_count, color = bin, fill = bin)) +
#     geom_density(alpha = 0.1) + 
#     labs(title = title_label,
#          x = "Log2 Transformed \n Normalized Read Counts",
#          y = "Density") +
#     theme_minimal()
# 
# 
# # Function to plot the signed log transformation
# signed_log_density_plot <- function(data_frame, title_label) {
#   data_frame %>%
#     mutate(normalized_read_count = as.numeric(normalized_read_count)) %>%
#     mutate(signed_log_read_count = sign(normalized_read_count) * log(abs(normalized_read_count) + 1)) %>% 
#     ggplot(aes(x = signed_log_read_count, color = bin, fill = bin)) +
#     geom_density(alpha = 0.1) + 
#     labs(title = title_label,
#          x = "Signed Log Transformed \n Normalized Read Counts",
#          y = "Density") +
#     theme_minimal()
# }
# # Function to plot the cube root transformation
# cube_root_density_plot <- function(data_frame, title_label) {
#   data_frame %>%
#     mutate(normalized_read_count = as.numeric(normalized_read_count)) %>%
#     mutate(cube_root_read_count = sign(normalized_read_count) * abs(normalized_read_count)^(1/3)) %>%
#     ggplot(aes(x = cube_root_read_count, color = bin, fill = bin)) +
#     geom_density(alpha = 0.1) + 
#     labs(title = title_label,
#          x = "Cube Root Transformed \n Normalized Read Counts",
#          y = "Density") +
#     theme_minimal()
# }
# 
# stmn2_log2_density_plot <- log2_density_plot(raw_counts_sgrna_stmn2, "STMN2 Screen")
# 
# stmn2_cuberoot_density_plot <- cube_root_density_plot(raw_counts_sgrna_stmn2, "STMN2 Screen")
# 
# 
# stmn2_signedlog_density_plot <- signed_log_density_plot(raw_counts_sgrna_stmn2, "STMN2 Screen")
# 
# 
# # Arrange plots side by side
# gridExtra::grid.arrange(stmn2_log2_density_plot, stmn2_cuberoot_density_plot, stmn2_signedlog_density_plot, ncol = 2, nrow = 2)
# gridExtra::grid.arrange(stmn2_cuberoot_density_plot, tdp_cuberoot_density_plot, iN3_cuberoot_density_plot, ipsc_cuberoot_density_plot, ncol = 2, nrow = 2)
# gridExtra::grid.arrange(stmn2_signedlog_density_plot, tdp_signedlog_density_plot, iN3_signedlog_density_plot, ipsc_signedlog_density_plot, ncol = 2, nrow = 2)
# 
# 

```

# SANDBOX 
```{r SANDBOX, echo=FALSE, eval=FALSE}
# Get column names
# col_names <- colnames(joined_df[-(1:2)])
# 
# # Generate all combinations of column names
# all_combinations <- lapply(1:length(col_names), function(x) combn(col_names, x, simplify = FALSE))
# 
# # Flatten the list of combinations
# all_combinations <- unlist(all_combinations, recursive = FALSE)
# 
# 
# # # Function to perform one-sample t-tests
# perform_t_test <- function(list_of_lists, df) {
#   col_name <- sample[[1]]
#   t_test_result <- t.test(df[[col_name]], mu = 0)
#   return(t_test_result)
# }
# 
# # Loop through each sample in the list and perform the t-test
# t_test_results <- lapply(differences_list, perform_t_test, df = tau_df_halo)
# 
# # Print the results
# names(t_test_results) <- sapply(samples, function(x) x[[1]])
# print(t_test_results)
```
